COCT Urban Mobility

### 2023-08-14 draft_v0_1.Rmd
Final draft typos and text sizes fixed. Dataset outside folder deleted. From Linux.

theme_doc <- theme_light() + 
  theme(text = element_text(color = "grey30",
                            size = 18),
        line = element_line(color = "grey30"),
        rect = element_rect(color = "grey30",
                            fill = "#FFFFF6"),
        panel.background = element_rect(fill = "#FFFFF6", colour = "grey70"),
        # axis.line = element_blank(),
        # axis.ticks = element_blank(),
        strip.background = element_rect(colour = "grey70",
                                        fill = "#FFFFF6"),
        strip.text = element_text(colour = "grey30"))

@@

appx_list$na_interpret_affected <- UpSetR::upset(appx_list$upset_data)

@@

```{r}
dta %>% 
  group_by(year_creation, month_creation) %>% 
  count(name = "total_ym") %>% 
  inner_join(dta %>% 
               group_by(year_creation, month_creation, directorate) %>% 
               count(name = "Nr_of_entries")) %>% 
  ungroup() %>% 
  mutate(Percentage_of_total = round(Nr_of_entries/total_ym*100, 2)) %>% 
  arrange(desc(Nr_of_entries)) %>% 
  ggplot(aes(month_creation, Percentage_of_total, 
             fill = reorder(directorate, Percentage_of_total)))+
  geom_col()+
  facet_wrap(~year_creation, scales = "free_x")+
  labs(title = "Figure 1",
    fill = "Directorate, ordered from smallest to largest contribution",
       x = "Month", y = "Percentage of total requests")

@@

```{r fig.height=8, fig.width=24}
map(unique(tmp$directorate),
    ~tmp %>% 
      filter(directorate == .x) %>% 
      arrange(desc(percentage)) %>% 
      slice(1:3)) %>% 
  bind_rows() %>% 
  ggplot(aes(percentage, official_suburb, fill = directorate))+
  geom_col()+
  facet_wrap(~directorate, scales = "free_y",
             nrow = 2)+
  scale_fill_manual(values = c(coct_cols, coct_cols))+
  labs(title = "Fig. 2",
    x = "Percentage")+
  theme(legend.position = "none",
        axis.title.y = element_blank())

@@

```{r focused5}

@@

```{r fig.height=10}
# nr over time, per suburb, missing suburbs excluded now

@@

```{r}
tmp <- dta_dir %>% 
  group_by(year_creation, month_creation) %>% 
  count(name = "total_month") %>% 
  inner_join(dta_dir %>% 
  group_by(year_creation, month_creation, official_suburb) %>% 
  count(name = "total_suburb")) %>% 
  ungroup() %>% 
    mutate(percentage = total_suburb/total_month*100) %>% 
    arrange(year_creation, month_creation, desc(percentage)) %>% 
  select(year_creation, month_creation, official_suburb,
         percentage) %>% 
    pivot_wider(names_from = month_creation, values_from = percentage) %>%  # add median across months
inner_join(dta_dir %>% 
             group_by(month_creation) %>% 
             count(name = "total_month") %>% 
             inner_join(dta_dir %>% 
  group_by(month_creation, official_suburb) %>% 
  count()) %>% 
  ungroup() %>% 
    mutate(percentage_suburb = n/total_month*100) %>% 
    group_by(official_suburb) %>% 
  summarise(med_percentage = median(percentage_suburb)) %>% 
  ungroup()) %>% 
  inner_join(tmp %>% 
               select(official_suburb)) %>% 
  distinct() %>% 
  arrange(desc(med_percentage))

@@

```{r}

@@

```{r}

@@

dta_dir %>% 
  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
  filter(creation_timestamp %in% x &
           official_suburb %in% top_ol) %>% 
  group_by(creation_timestamp, official_suburb, code) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  filter(n != 1) %>% 
  kable(caption = "Outier dates", format = "simple")

@@

```{r}
tmp <- dta_dir %>% 
  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
  filter(!creation_timestamp %in% x)

@@

```{r}
# explore codes for top 
select_suburbs <- c("MARCONI BEAM", "TABLE VIEW", "MILNERTON")

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}
x <- time_dta_list$all_times %>% 
  filter(official_suburb == "ALL COMBINED" &
           code == "All codes combined") %>% 
  mutate(perc = round(x/n*100,2)) %>% 
  pull(perc)

@@

```{r}
# filter by < 24 hrs

@@

```{r}
# split by 24 hours

@@

```{r}

@@

```{r}
# split by codes
# codes with > 5 requests, per suburb
# show top 5 per code

@@

```{r}

@@

```{r fig.height=10}
time_dta_list$all_times %>% 
  filter(official_suburb %in% sub_txt &
           code != code_txt &
           n >= n_cutoff) %>% 
  pivot_longer(c(med_time_hours, q80_time_hours),
               names_to = "quantile", 
               values_to = "hours") %>% 
  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
   geom_line(colour = "grey70")+
  geom_point(size = 3)+
  facet_wrap(~official_suburb)+
   scale_colour_manual(values = coct_cols)+
  theme(axis.title.y = element_blank())+
  labs(title = "Figure 8", caption = "X-axis shows log base 10 transformed hours,
       so that 1 denotes 10 hours, 2, 100 hours, etc. Median
       time is shown in pink and 80% percentile in green.")

@@

```{r}

@@

```{r}

@@

```{r}
# traffic congestion over time

@@

x <- sort(tmp$n, decreasing = TRUE)
x <- x[6] # for top 5

tmp %>% 
mutate(day_of_week = wday(creation_timestamp, label=TRUE),
       label = case_when(n > x ~ 
                           paste(day_of_week, creation_timestamp),
                         TRUE ~ ""),
       weekend = case_when(day_of_week %in% c("Sat", "Sun") ~ "weekend",
                           TRUE ~ "week")) %>%

 ggplot(aes(creation_timestamp, n, group = code,
            label = label, colour = weekend))+
 geom_line(colour = "grey70")  +
 geom_point(size = 3)+
   geom_text()+
  scale_color_manual(values = coct_cols)+
  theme(legend.title = element_blank())+
  labs(title = "Figure 9",
       caption = "X-axis shows dates. Dots are coloured by week day (pink) or weekend day (green).
       Outlier points are labelled with dates.")

@@

```{r fig.height = 16}
dta_dir %>% 
 mutate(create_date = as_date(creation_timestamp)) %>% 
 filter(code == code_txt) %>% 
 group_by(code,create_date) %>% 
 count() %>% 
 ungroup() %>% 
  filter(n > x) %>% 
  select(-n) %>% 
  inner_join(dta_dir) %>% 
  group_by(create_date,official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
  geom_col()+
  scale_fill_manual(values = coct_cols)+
  labs(fill = "Top 5 Dates")+
  theme(axis.title.y = element_blank())+
  labs(title = "Figure 10",
       caption = "Each bar show number of requests, with different colours for each date's contribution.")

@@

```{r}

@@

```{r}
 # suburbs and days of week

@@

sz <- 3
 dta_dir %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(day_of_week, official_suburb) %>% 
 count() %>% 
   mutate(label = case_when(n >= x ~ official_suburb,
                          TRUE ~ "")) %>%
   ggplot(aes(day_of_week, n, label = label))+
   geom_jitter(alpha = 0.6, width = 0.1, colour = coct_cols[1])+
   ggrepel::geom_text_repel(alpha = 0.5, size = sz,
                            max.overlaps = 50,
                            force = 0.1)+
   labs(title = "Figure 12",
        caption = "Each dot represents a suburb, only ones with
        high numbers are labelled.")

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

>>>>

theme_doc <- theme_light() + 
  theme(text = element_text(color = "grey30",
                            size = 19),
        line = element_line(color = "grey30"),
        rect = element_rect(color = "grey30",
                            fill = "#FFFFF2"),
        panel.background = element_rect(fill = "#FFFFF2", colour = "grey70"),
        # axis.line = element_blank(),
        # axis.ticks = element_blank(),
        strip.background = element_rect(colour = "grey70",
                                        fill = "#FFFFF2"),
        strip.text = element_text(colour = "grey30"))

@@

appx_list$na_interpret_affected <- UpSetR::upset(appx_list$upset_data, text.scale = 3)

@@

```{r overview2}
dta %>% 
  group_by(year_creation, month_creation) %>% 
  count(name = "total_ym") %>% 
  inner_join(dta %>% 
               group_by(year_creation, month_creation, directorate) %>% 
               count(name = "Nr_of_entries")) %>% 
  ungroup() %>% 
  mutate(Percentage_of_total = round(Nr_of_entries/total_ym*100, 2)) %>% 
  arrange(desc(Nr_of_entries)) %>% 
  ggplot(aes(month_creation, Percentage_of_total, 
             fill = reorder(directorate, Percentage_of_total)))+
  geom_col()+
  facet_wrap(~year_creation, scales = "free_x")+
  labs(title = "Figure 1",
    fill = "Directorate, ordered from smallest to largest contribution",
       x = "Month", y = "Percentage of total requests")

@@

```{r overview3,fig.height=8, fig.width=24}
map(unique(tmp$directorate),
    ~tmp %>% 
      filter(directorate == .x) %>% 
      arrange(desc(percentage)) %>% 
      slice(1:3)) %>% 
  bind_rows() %>% 
  ggplot(aes(percentage, official_suburb, fill = directorate))+
  geom_col()+
  facet_wrap(~directorate, scales = "free_y",
             nrow = 2)+
  scale_fill_manual(values = c(coct_cols, coct_cols))+
  labs(title = "Fig. 2",
    x = "Percentage")+
  theme(legend.position = "none",
        axis.title.y = element_blank())

@@

```{r focused3}

@@

```{r focused5, fig.height=10}
# nr over time, per suburb, missing suburbs excluded now

@@

```{r focused6}
tmp <- dta_dir %>% 
  group_by(year_creation, month_creation) %>% 
  count(name = "total_month") %>% 
  inner_join(dta_dir %>% 
  group_by(year_creation, month_creation, official_suburb) %>% 
  count(name = "total_suburb")) %>% 
  ungroup() %>% 
    mutate(percentage = total_suburb/total_month*100) %>% 
    arrange(year_creation, month_creation, desc(percentage)) %>% 
  select(year_creation, month_creation, official_suburb,
         percentage) %>% 
    pivot_wider(names_from = month_creation, values_from = percentage) %>%  # add median across months
inner_join(dta_dir %>% 
             group_by(month_creation) %>% 
             count(name = "total_month") %>% 
             inner_join(dta_dir %>% 
  group_by(month_creation, official_suburb) %>% 
  count()) %>% 
  ungroup() %>% 
    mutate(percentage_suburb = n/total_month*100) %>% 
    group_by(official_suburb) %>% 
  summarise(med_percentage = median(percentage_suburb)) %>% 
  ungroup()) %>% 
  inner_join(tmp %>% 
               select(official_suburb)) %>% 
  distinct() %>% 
  arrange(desc(med_percentage))

@@

```{r focused7}

@@

```{r focused8}

@@

dta_dir %>% 
  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
  filter(creation_timestamp %in% x &
           official_suburb %in% top_ol) %>% 
  group_by(creation_timestamp, official_suburb, code) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  filter(n != 1) %>% 
  kable(caption = "Outlier dates", format = "simple")

@@

```{r focused9}
tmp <- dta_dir %>% 
  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
  filter(!creation_timestamp %in% x)

@@

```{r focused10}
# explore codes for top 
select_suburbs <- c("MARCONI BEAM", "TABLE VIEW", "MILNERTON")

@@

```{r times1}

@@

```{r times2}

@@

```{r times3}

@@

```{r times4}
x <- time_dta_list$all_times %>% 
  filter(official_suburb == "ALL COMBINED" &
           code == "All codes combined") %>% 
  mutate(perc = round(x/n*100,2)) %>% 
  pull(perc)

@@

```{r times5}
# filter by < 24 hrs

@@

```{r times6}
# split by 24 hours

@@

```{r times7}

@@

```{r times8}
# split by codes
# codes with > 5 requests, per suburb
# show top 5 per code

@@

```{r times9}

@@

```{r times10, fig.height=10}
time_dta_list$all_times %>% 
  filter(official_suburb %in% sub_txt &
           code != code_txt &
           n >= n_cutoff) %>% 
  pivot_longer(c(med_time_hours, q80_time_hours),
               names_to = "quantile", 
               values_to = "hours") %>% 
  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
   geom_line(colour = "grey70")+
  geom_point(size = 3)+
  facet_wrap(~official_suburb)+
   scale_colour_manual(values = coct_cols)+
  theme(axis.title.y = element_blank())+
  labs(title = "Figure 8", caption = "X-axis shows log base 10 transformed hours,
       so that 1 denotes 10 hours, 2, 100 hours, etc. Median
       time is shown in pink and 80% percentile in green.")

@@

```{r times11}

@@

```{r dash}

@@

```{r insights1}
# traffic congestion over time

@@

x <- sort(tmp$n, decreasing = TRUE)
x <- x[6] # for top 5
sz <- 5
tmp %>% 
mutate(day_of_week = wday(creation_timestamp, label=TRUE),
       label = case_when(n > x ~ 
                           paste(day_of_week, creation_timestamp),
                         TRUE ~ ""),
       weekend = case_when(day_of_week %in% c("Sat", "Sun") ~ "weekend",
                           TRUE ~ "week")) %>%

 ggplot(aes(creation_timestamp, n, group = code,
            label = label, colour = weekend))+
 geom_line(colour = "grey70")  +
 geom_point(size = 3)+
   geom_text(size = sz)+
  scale_color_manual(values = coct_cols)+
  theme(legend.title = element_blank())+
  labs(title = "Figure 9",
       caption = "X-axis shows dates. Dots are coloured by week day (pink) or weekend day (green).
       Outlier points are labelled with dates.")

@@

```{r insights2,fig.height = 16}
dta_dir %>% 
 mutate(create_date = as_date(creation_timestamp)) %>% 
 filter(code == code_txt) %>% 
 group_by(code,create_date) %>% 
 count() %>% 
 ungroup() %>% 
  filter(n > x) %>% 
  select(-n) %>% 
  inner_join(dta_dir) %>% 
  group_by(create_date,official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
  geom_col()+
  scale_fill_manual(values = coct_cols)+
  labs(fill = "Top 5 Dates")+
  theme(axis.title.y = element_blank())+
  labs(title = "Figure 10",
       caption = "Each bar show number of requests, with different colours for each date's contribution.")

@@

```{r insights3}

@@

```{r insights4}
 # suburbs and days of week

@@

sz <- 5
 dta_dir %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(day_of_week, official_suburb) %>% 
 count() %>% 
   mutate(label = case_when(n >= x ~ official_suburb,
                          TRUE ~ "")) %>%
   ggplot(aes(day_of_week, n, label = label))+
   geom_jitter(alpha = 0.6, width = 0.1, colour = coct_cols[1])+
   ggrepel::geom_text_repel(alpha = 0.5, size = sz,
                            max.overlaps = 50,
                            force = 0.1)+
   labs(title = "Figure 12",
        caption = "Each dot represents a suburb, only ones with
        high numbers are labelled.")

@@

```{r appx1}

@@

```{r appx2}

@@

```{r appx3}

@@

```{r appx4}

@@

```{r appx5}

@@

```{r appx6}
